Identifying Software Performance Changes Across Variants and Versions

### 2020-08-11 paper.tex
amended introduction, now focusing on retrospective performance analysis

	%\paragraph{\bfseries Why has it not been solved before?}
	Our goal is to identify changes in the performance histories of configurable software systems and pinpoint them to a specific option or interaction. For example, a patch of a certain feature is likely to affect only configurations where such feature is selected.
	In essence, we face a huge combinatorial problem for performance-change detection: code changes across software versions and behavioral changes across software variants.
	Our problem arises along two dimensions: time (versions) and configuration space (variants).
	\emph{First}, we aim at finding performance changes in software performance from one version to another.
	The detection of such changes is referred to as \emph{change point detection} and is a recurring theme in temporal analysis.
	The main limiting factor is that exhaustive measurements across the configuration space are neither available nor feasible~\cite{white_selecting_2009}.
	Change point detection techniques with both exhaustive measurement~\cite{cityIdentifying2014,daly_industry_2020} as well as limited data availability~\cite{sandoval_alcocer_learning_2016,alcocer_prioritizing_2020,huang_performance_2014,muhlbauer_accurate_2019} have been successfully applied to identify performance changes.
	\emph{Second}, we aim at associating changes in performance (from one version to another) with particular configuration options or interactions among them.
	There is substantial work  regarding a related problem, which is estimating the influence of individual options on performance~\cite{siegmundPredictingPerformanceAutomated2012,siegmundPerformanceinfluenceModelsHighly2015,sarkarCostEfficientSamplingPerformance,haDeepPerf2019}. Instead of estimating the influence of options and interactions on performance, we want to know: Which option or interaction is \emph{responsible} for a particular change point?

@@

	Our main idea is as follows: We address the configuration complexity of this problem by selecting representative sample sets of configurations.
	Next, we uniformly sample a constant number of commits for each configuration and conduct respective performance measurements.
	Based on these measurement data, we assemble a prediction model.
	For each configuration, we estimate the likelihood of each commit being a change point, i.e. that performance of some configurations changes abruptly compared to the previous commit.
	Next, we leverage similarities in the performance histories of configurations that share common options.
	Since we sample the commits for each configuration independently, for each change point, we obtain many estimations using measurements of different commits.
	Overall, this allows us to obtain more accurate estimations of performance-changing commits with tractable effort.
	Based on a mapping from configurations to predicted change point probabilities for each configuration and commit, we derive the options responsible for each particular change point: configuration-dependent change points.
	%\paragraph{\bfseries What are our contributions?}
	In summary, we offer the following contributions:
	\begin{itemize}
		\item A {\color{red}novel perspective and} technique to effectively identify shifts in performance of configurable software systems. We are able to pinpoint causative commits and affected configuration options with high accuracy and tractable effort.
		\item A feasibility demonstration of our approach by implementing an adaptive learning strategy to obtain accurate estimations with acceptable measurement cost.
		\item An evaluation using both synthetic and real-world performance data from three configurable software systems. Synthetic data lets us assess our model and approach conceptually and at scale, whereas we are able to assess practicality with real-world data.
		\item A companion website\footnote{\url{https://github.com/AI-4-SE/Changepoints-Across-Variants-And-Versions}} providing supplementary material including a reference implementation of our approach, performance measurement data, and additional visualizations.
	\end{itemize}

>>>>

	%\paragraph{\bfseries Why has it not been solved before?}
	%Our goal is to identify changes in the performance histories of configurable software systems and pinpoint them to a specific option or interaction. For example, a patch of a certain feature is likely to affect only configurations where such feature is selected.
	%In essence, we face a huge combinatorial problem for performance-change detection: code changes across software versions and behavioral changes across software variants.

	{\color{red}Our goal is to enable the \emph{retrospective} detection of changes in the performance of configurable software systems and pinpoint them to a specific option or interaction. For example, a patch of a certain feature is likely to affect only configurations where such feature is selected. Many software systems are configurable, but have no performance regression testing routine set in place. To uncover performance deficiencies that emerged from revisions in the past development history, it is infeasible to test each and every commit and configuration.} In essence, we face a huge combinatorial problem for performance-change detection: code changes across software versions and behavioral changes across software variants.
	Our problem arises along two dimensions: time (versions) and configuration space (variants).
	\emph{First}, we aim at finding performance changes in software performance from one version to another.
	The detection of such changes is referred to as \emph{change point detection} and is a recurring theme in temporal analysis.
	The main limiting factor is that exhaustive measurements across the configuration space are neither available nor feasible~\cite{white_selecting_2009}.
	Change point detection techniques with both exhaustive measurement~\cite{cityIdentifying2014,daly_industry_2020} as well as limited data availability~\cite{sandoval_alcocer_learning_2016,alcocer_prioritizing_2020,huang_performance_2014,muhlbauer_accurate_2019} have been successfully applied to identify performance changes.
	\emph{Second}, we aim at associating changes in performance (from one version to another) with particular configuration options or interactions among them.
	There is substantial work  regarding a related problem, which is estimating the influence of individual options on performance~\cite{siegmundPredictingPerformanceAutomated2012,siegmundPerformanceinfluenceModelsHighly2015,sarkarCostEfficientSamplingPerformance,haDeepPerf2019}. Instead of estimating the influence of options and interactions on performance, we want to know: Which option or interaction is \emph{responsible} for a particular change point?

@@

	Our main idea is as follows: We address the configuration complexity of this problem by selecting representative sample sets of configurations.
	Next, we uniformly sample a constant number of commits for each configuration and conduct respective performance measurements.
	Based on these measurement data, we assemble a prediction model.
	For each configuration, we estimate the likelihood of each commit being a change point, i.e. that performance of some configurations changes abruptly compared to the previous commit.
	Next, we leverage similarities in the performance histories of configurations that share common options.
	Since we sample the commits for each configuration independently, for each change point, we obtain many estimations using measurements of different commits.
	Overall, this allows us to obtain more accurate estimations of performance-changing commits with tractable effort.
	Based on a mapping from configurations to predicted change point probabilities for each configuration and commit, we derive the options responsible for each particular change point: configuration-dependent change points. {\color{red}g}
	In summary, we offer the following contributions:
	\begin{itemize}
		\item A {\color{red}novel perspective and} technique to effectively identify shifts in performance of configurable software systems. We are able to pinpoint causative commits and affected configuration options with high accuracy and tractable effort.
		\item A feasibility demonstration of our approach by implementing an adaptive learning strategy to obtain accurate estimations with acceptable measurement cost.
		\item An evaluation using both synthetic and real-world performance data from three configurable software systems. Synthetic data lets us assess our model and approach conceptually and at scale, whereas we are able to assess practicality with real-world data.
		\item A companion website\footnote{\url{https://github.com/AI-4-SE/Changepoints-Across-Variants-And-Versions}} providing supplementary material including a reference implementation of our approach, performance measurement data, and additional visualizations.
	\end{itemize}
